---
title: "Nov27_Practice"
author: "Brandie Quarles"
date: "2023-11-20"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(lmerTest)
```


We will use the ChickWeight data set. It is a built in dataset For info:

```{r}
?ChickWeight
#Weight versus age of chicks on different diets
```

To use it:

```{r}
data(ChickWeight)
# then it is an object in your environment
summary(ChickWeight)

# Or if you have skimr installed
skimr::skim(ChickWeight) #nicer summary

head(ChickWeight)
#this is an example of repeated measures, each chick measured at different time points 
#also each chick is nested within a diet (only 1 diet per chick)
```

Plot the data
```{r}
ChickWeight %>%
  ggplot(aes(x=Time, y=weight)) +
  geom_line(aes(group=Chick), alpha=.2) +
  geom_smooth() +
  facet_wrap(~Diet)
```


1.  Use rsample to create training and test sets from ChickWeight. Think
    about how you want this to be split. hint:: you may need to also use
    a function from Chapter 3. message me if you are stuck.
```{r}
set.seed(510)
chicks_split_simple <- initial_split(ChickWeight, prop=0.80)
chicks_split_simple
chicks_train_simple <- training(chicks_split_simple)
chicks_test_simple  <-  testing(chicks_split_simple)
#this isn't good b/c it splits up data from a given chick 
```

Simple + Strata by weight
```{r}
ggplot(ChickWeight, aes(x = weight)) + 
  geom_histogram(bins = 50, col= "white") #data very right skewed by weight
#want stratified sampling 

set.seed(511)
chicks_split_strata <- initial_split(ChickWeight, prop=0.80, strata=weight)
chicks_split_strata
chicks_train_strata <- training(chicks_split_strata)
chicks_test_strata  <-  testing(chicks_split_strata)
#stratifying by weight is good, but still need to fix the chicks problem
```

I need it to put different chicks in different sets, but keep all of a single chick's info together:
```{r}
set.seed(512)
chicks_split_chick <- 
  ChickWeight %>% 
  group_nest(Chick)  %>% 
  initial_split(prop=0.80)
chicks_split_chick
chicks_train_chick <- training(chicks_split_chick)
chicks_test_chick  <-  testing(chicks_split_chick)

#won't let me also use a strata argument with this syntax
```

Julin's Version:
```{r}
#want chicks from different diets in both sets 
chicken.gr <- ChickWeight %>% group_by(Chick, Diet) %>% nest()

chick.split <- chicken.gr %>% 
  rsample::initial_split(prop = .8, strata = Diet)

chick.split
#alternatively could use group_initial_split()

chick.train <- training(chick.split) %>% unnest(cols = data)
chick.test <- testing(chick.split) %>% unnest(cols = data)
chick.train

ChickWeight %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
chick.train %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
chick.test %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
```

```{r}
#John's version
grouped_split_chick <- ChickWeight %>%
  group_nest(Chick, Diet) %>%
  initial_split(prop = 0.80, strata = Diet)

chick_train <- training(grouped_split_chick) %>% unnest(data)
chick_test<- testing(grouped_split_chick) %>% unnest(data)
```



2.  Fit an lm to the ChickWeight (training) data set, modelling weight
    as a function of diet and time (and their interaction?), using
    parsnip tools. This really should be a mixed-effects model, so see
    if you can do that instead (e.g. with lme4 or stan as the engine).
```{r}
#regular linear model
lm_model <- 
  linear_reg() %>% 
  set_engine("lm")

chicks_train_chick_ungrouped <- chicks_train_chick %>% unnest(data)
chicks_train_chick_ungrouped

lm_fit <- 
  lm_model %>% 
  fit(weight ~ Diet + Time, data = chicks_train_chick_ungrouped)

lm_fit %>% extract_fit_engine()

model_res <- 
  lm_fit %>% 
  extract_fit_engine() %>% 
  summary()
model_res

param_est <- coef(model_res)
class(param_est)

param_est

tidy(lm_fit) 
```

```{r}
#mixed-effects model
library(multilevelmod)
show_engines('linear_reg')
mixed_lm_model <-
  linear_reg() %>% 
  set_engine("lmer")

unique(chicks_train_chick_ungrouped$Time) #time should be nested within Chick which is nested within diet
unique(chicks_train_chick_ungrouped$Diet)
lm_mixed_fit <- 
  mixed_lm_model %>% 
  fit(weight ~ Diet + (1|Diet/Time), data = chicks_train_chick_ungrouped) #I'm pretty sure this is the wrong syntax

#best version
#fit(weight ~ Time + Diet +  (1|Chick)) # works b/c each chick has an individual id
#1| says fit a different intercept for each chick 

#another option to explore 
#fit(weight ~ Time + Diet +  (1 + Time|Chick)) #different slope for each time by chick and different intercepts 

#interaction term
#fit(weight ~ Time*Diet + (1|Chick))

lm_mixed_fit %>% extract_fit_engine()

model_res <- 
  lm_mixed_fit %>% 
  extract_fit_engine() %>% 
  summary()
model_res

param_est <- coef(model_res)
class(param_est)

param_est

#tidy(lm_mixed_fit) #Error: No tidy method for objects of class lmerMod
#Need: 
library(broom.mixed)

#To compare models use #bind_rows(glance(fit1), glance(fit2))
#tidy(fit2)
#or anova(extract_fit_engine(fit1), extract_fit_engine(fit2))


```


3.  Use your model to predict weight in your test set chicks. (using
    parsnip tools)
```{r}
chicks_test_chick_ungrouped <- chicks_test_chick %>% unnest(data)
chicks_test_chick_ungrouped

predict(lm_fit, new_data = chicks_test_chick_ungrouped)

pred_obs_data <- chicks_test_chick_ungrouped %>% 
  select(weight) %>% 
  bind_cols(predict(lm_fit, chicks_test_chick_ungrouped)) %>% 
  # Add 95% prediction intervals to the results:
  bind_cols(predict(lm_fit, chicks_test_chick_ungrouped, type = "pred_int")) 
pred_obs_data

#If you have multiple models can ocmbine them into a tibble
#tibble(name=ls(.GlobalEnv, pattern=".*_fit.*"), fit=) #see Julin's code 
#prediction via parsnip may not be able to take into account random effects from mixed effects models
#can use predict.merMod outside of parsnip context #can make predictions for new block or new chick for example 
```

4.  Plot predicted vs observed in your test data set.
```{r}
pred_obs_data %>% ggplot(aes(weight, .pred)) + geom_point()

#ideal = different colors for different diets 

#Julin also plotted time on x and weihgt on y w/ pred and obs as different types of lines 
fits <- tibble(name=ls(.GlobalEnv, pattern=".*_fit.*"),
               fit = map(name, get))
fits

fits <- fits %>% mutate(pred=map(fit, predict, new_data=chick.test),
                        pred=map(pred, bind_cols, chick.test)) # add observed

fits %>% select(-fit) %>% unnest(pred)

fits %>% select(-fit) %>% unnest(pred) %>%
  ggplot(aes(x=weight, y=.pred, color=Diet)) +
  scale_color_brewer(type = "qual", palette = "Set2") +
  geom_point() +
  facet_wrap(~name) +
  coord_equal()

fits %>% select(-fit) %>% unnest(pred) %>% 
  rename(obs = weight, model=name) %>%
  pivot_longer(cols=c(obs, .pred)) %>%
  ggplot(aes(x=Time, y=value, color=Diet, lty=name)) +
  scale_color_brewer(type = "qual", palette = "Set1") +
  geom_smooth(se = FALSE) +
  facet_wrap(~model)
```


5.  Optional: recreate an Ames Neighborhood plot.

    1.  Tutorials are at <https://ggplot2tutor.com/tutorials/streetmaps>
        <https://joshuamccrain.com/tutorials/maps/streets_tutorial.html>
